Using traffic analysis to identify The Second Generation Onion Router

### 2011-02-27 thesis/experiment.tex
Split out into different sections, more rewriting as per suggestions.

  OpenSSH is a freely available suite of connectivity tools
  \parencite{:2010zr,:ve} that supports remote command execution
  \parencite{Tucker:2010ly, :2010zr}. The client component of OpenSSH is
  installed by default on the Ubuntu operating system and the server component
  is easily installed using the apt-get package management system.

@@

  RSpec is a testing framework for Ruby which allows the creation of examples in
  an easy to read Domain Specific Language (DSL). It is well integrated with
  Selenium which makes it ideal for running large suites of Selenium tests.

@@

Four identical desktop computer systems were prepared and arranged as shown
in figure \ref{physical-setup}.

\begin{figure}[H]
  \centering\includegraphics[scale=0.7]{physical-setup}
  \caption{Physical Setup}
  \label{physical-setup}
\end{figure}

@@

\begin{tabular}{lrr}
  \toprule
  Role & Operating System Version & IP Address\\
  \midrule
  Client Host & Desktop & 192.168.0.100\\
  Server Host & Desktop & 192.168.0.101\\
  Tor Host & Desktop & 192.168.0.102\\
  \midrule
  Sniffing System & Server & 192.168.0.200\\
  \midrule
  Client Guest & Desktop & 192.168.0.16\\
  Server Guest & Server & 192.168.0.17\\
  Tor Network Guest & Server & 192.168.0.18\\
  \bottomrule
  \label{table:hosts}
\end{tabular}

Figure \ref{network-diagram} shows the placement of applications and relevant
data flows when running the simulation in HTTPS mode.

\begin{figure}[H]
  \centering\includegraphics[width=\linewidth]{network-diagram}
  \caption{Network Diagram: Phase 1}
  \label{network-diagram}
\end{figure}

Figure \ref{network-diagram-with-tor} shows the network configuration when
running the experiment in HTTP over Tor mode.

\begin{figure}[H]
  \centering\includegraphics[width=\linewidth]{network-diagram-with-tor}
  \caption{Network Diagram: Phase 2}
  \label{network-diagram-with-tor}
\end{figure}

Figure \ref{network-diagram-with-tor-encrypted} shows the network configuration when
running the experiment in HTTPS over Tor mode.

\begin{figure}[H]
  \centering\includegraphics[width=\linewidth]{network-diagram-with-tor-encrypted}
  \caption{Network Diagram: Phase 3}
  \label{network-diagram-with-tor-encrypted}
\end{figure}

@@

All systems in the experiment use a variation of the Ubuntu Linux operating
system. The installation process followed is detailed below:

\begin{enumerate*}
  \item Configure system BIOS to boot from CD-ROM
  \item Insert Ubuntu operating system installation CD
  \item Select the language you wish to use for installation (English) and click
    'Forward'
  \item Select the region (Australia/Perth) and click 'Forward'
  \item Select an appropriate keymap (USA) and click 'Forward'
  \item Select 'Erase and use the entire disk' and click 'Forward'
  \item Enter the user name (John Barker), select 'Log in automatically' and click 'Forward'
\end{enumerate*}

At this point the installation will begin and should take about 35 minutes.
'Log in automatically' is selected to ensure that the experiment will continue
even if power is lost. Once installation is complete remove the CD and reboot
the system. Once booted the system is ready for further configuration.

@@

An init.d script was created to execute the tcpdump program with a specific
filter string after the system had powered on. All capture files were limited to
1mb in size and put into a time stamped directory. To ensure that only relevant
packets were collected the filter string was tuned for each phase of the
experiment. The filter string for each phase is shown below.

@@

\begin{tabular}{lrr}
  \toprule
  Phase & Size (bytes) & Number of sessions\\
  \midrule
  HTTPS & 2326328 & 236659\\
  HTTP over Tor & 2869624 & 168876\\
  HTTPS over Tor & 1591112 & 95203\\
  \bottomrule
  \label{table:datasets}
\end{tabular}

The data captured was in the form of 1mb capture files which were recombined
with mergecap and processed by NetAI to produce ARFF format files for use by
Weka. Each data set were given the nominal classification 'http',
'http\_over\_tor' and 'https\_over\_tor', all other nominal attributes were
removed.

>>>>

  OpenSSH is a freely available suite of connectivity tools
  \parencite{:2010zr,:ve} that supports remote command execution
  \parencite{Tucker:2010ly}. The client component of OpenSSH is installed by
  default on the Ubuntu operating system and the server component is easily
  installed using the apt-get package management system.

@@

  RSpec is a behaviour driven testing framework for Ruby which allows the
  creation of examples in an easy to read Domain Specific Language (DSL)
  \parencite{Chelimsky:2010vn}. It is well integrated with Selenium which makes
  it ideal for running large suites of Selenium tests.

@@

Four identical desktop computer systems were prepared and arranged as shown
in figure \ref{fig:physical-setup}.

\begin{figure}[H]
  \centering\includegraphics[scale=0.7]{physical-setup}
  \caption{Physical Setup}
  \label{fig:physical-setup}
\end{figure}

@@

\begin{table}[H]
  \begin{tabular}{lrr}
    \toprule
    Role & Operating System Version & IP Address\\
    \midrule
    Client Host & Desktop & 192.168.0.100\\
    Server Host & Desktop & 192.168.0.101\\
    Tor Host & Desktop & 192.168.0.102\\
    \midrule
    Sniffing System & Server & 192.168.0.200\\
    \midrule
    Client Guest & Desktop & 192.168.0.16\\
    Server Guest & Server & 192.168.0.17\\
    Tor Network Guest & Server & 192.168.0.18\\
    \bottomrule
  \end{tabular}
  \caption{System Roles}
  \label{table:hosts}
\end{table}

Figure \ref{fig:network-diagram} shows the placement of applications and relevant
data flows when running the simulation in HTTPS mode.

\begin{figure}[H]
  \centering\includegraphics[width=\linewidth]{network-diagram}
  \caption{Network Diagram: Phase 1}
  \label{fig:network-diagram}
\end{figure}

Figure \ref{fig:network-diagram-with-tor} shows the network configuration when
running the experiment in HTTP over Tor mode.

\begin{figure}[H]
  \centering\includegraphics[width=\linewidth]{network-diagram-with-tor}
  \caption{Network Diagram: Phase 2}
  \label{fig:network-diagram-with-tor}
\end{figure}

Figure \ref{fig:network-diagram-with-tor-encrypted} shows the network configuration when
running the experiment in HTTPS over Tor mode.

\begin{figure}[H]
  \centering\includegraphics[width=\linewidth]{network-diagram-with-tor-encrypted}
  \caption{Network Diagram: Phase 3}
  \label{fig:network-diagram-with-tor-encrypted}
\end{figure}

@@

All systems in the experiment use a variation of the Ubuntu Linux operating
system. Table \ref{table:config-options} specifies the options supplied when
installing each operating system for the experiment.

\begin{table}[H]
  \begin{tabular}{lr}
    \toprule
    Prompt & Value\\
    \midrule
    Where are you? & Australia/Perth\\
    Which layout is most similar to your keyboard? & USA\\
    How do you want to partition the disk? & Erase and use the entire disk\\
    What is your name? & John Barker\\
    What name do you want to use to log in? & john\\
    \bottomrule
  \end{tabular}
  \caption{Ubuntu Installation Parameters}
  \label{table:config-options}
\end{table}

@@

An init.d script was created to execute the tcpdump program with a specific
filter string after the system had powered on. All capture files were limited to
1MB in size and put into a time stamped directory. To ensure that only relevant
packets were collected the filter string was tuned for each phase of the
experiment. The filter string for each phase is shown below.

@@

\begin{table}[H]
  \begin{tabular}{lrr}
    \toprule
    Phase & Size in bytes & Number of sessions\\
    \midrule
    HTTPS & 2326328 & 236659\\
    HTTP over Tor & 2869624 & 168876\\
    HTTPS over Tor & 1591112 & 95203\\
    \bottomrule
    \label{table:datasets}
  \end{tabular}
  \caption{Captured Data}
\end{table}

The data captured was in the form of 1MB capture files which were recombined
with mergecap and processed by NetAI to produce ARFF format files for use by
Weka. Each data set were given the nominal classification 'http',
'http\_over\_tor' and 'https\_over\_tor', all other nominal attributes were
removed.

### 2011-02-27 thesis/litreview.tex
Split out into different sections, more rewriting as per suggestions.

\documentclass{ecuthesis}

% For graphics
\usepackage{graphicx}

% Bibliography
\bibliography{fullrefs,websites,patents,papers}

% Setup the variables for the title page
\title{Using traffic analysis to identify The Second Generation Onion Router}
\author{John Barker}
\department{School of Computer and Security Science}
\school{Edith Cowan University}
\degree{Computer Science (Honours)}
\email{jebarker@our.ecu.edu.au}
\studentid{0991300}
\supervisor{Dr Andrew Woodward}
\supervisor{Mr Peter Hannay}
\supervisor{Mr Patryk Szewczyk}

% Begin document
\begin{document}
\maketitle

\tableofcontents
\listoftables
\listoffigures

\begin{abstract}
\thispagestyle{empty}

Anonymous networks provide security for users by obfuscating messages with
encryption and hiding communications amongst cover traffic provided by other
network participants. The traditional goal of academic research into these
networks has been attacks that aim to uncover the identity of network users.
But the success of an anonymous network relies not only on it's technical
capabilities, but on adoption by a large enough user base to provide adequate
cover traffic. If anonymous network nodes can be identified, the users
can be harassed, discouraging participation. Tor is an example of widely used
anonymous network which uses a form of Onion Routing to provide low latency
anonymous communications. This paper demonstrates that traffic from a simulated
Tor network can be distinguished from regular encrypted traffic, suggesting that
real world Tor users may be vulnerable to the same analysis.

\end{abstract}

\chapter{Introduction}

\section{Background}

The Internet has revolutionised the political sphere in providing a platform
for the publication of speech to a far greater audience than was ever available
before digital communications systems \parencite{Bonchek:1997p3455}. The
publication of less desirable political speech, criticism or challenging ideas
still carries with it great risk, with numerous publications leading to arrest.
This has spawned headlines such as \textcite{website:egypt-arrests},
\citetitle{website:china-yahoo-torture}
\parencite{website:china-yahoo-torture},
\citetitle{website:vietnam-bloggers-arrested}
\parencite{website:vietnam-bloggers-arrested},
\textcite{website:iran-bloggers-arrested} and
\citetitle{website:blogger-arrests} \parencite{website:blogger-arrests}. The
Internet, like conventional media, is still vulnerable to censorship by
oppressive governments and malicious attackers
\parencite{Crandall:2007p6165,Karlin:2009p6166}.

In response to these threats, a number of systems have been proposed which use
various techniques, primarily cryptography and covering traffic, to provide
censorship resistance and anonymity. The Second Generation Onion Router (Tor),
is an anonymous network which uses a variant of Onion Routing, to provide low
latency anonymous communications to a wide variety of Internet communications
protocols \parencite{Dingledine:2004p314}. It also aims to provide resistance
against censorship based blocking attempts \parencite{Dingledine:2008p1542}.

Although Tor succeeds to a certain extent in these goals, it may be possible to
use traffic analysis techniques to automatically identify Tor connections and
target them for attacks. Since Tor relies on large numbers of users for the
protections it provides, attacks that affect the reliability of the network can
discourage participation, making it ineffective. This paper proposes a traffic
analysis methodology for identifying Tor traffic which may be used for such
attacks in the future.

\section{Significance}

Tor is one of the most successful and widely deployed anonymous networks, with
around one thousand, eight hundred users at the time of writing
\parencite{website:tor-anonymity-online}. While it has been the subject of many
traffic analysis attacks, the primary objective has historically been the
attempt to reveal participant identities: ``The principal objective of an
adversary attacking an anonymous communication system is to link the initiators
of connections with their respective communication partners and vice versa''
\parencite[3]{Murdoch:2005p325}.

The traffic analysis techniques proposed will attempt to classify network
streams belonging to Tor users to distinguish them from regular encrypted
communications. This research may provide a greater understanding of just
how much anonymity a user of the Tor network is provided and inspire
the development of countermeasures against analysis.

\section{Research Questions}

Given a motivated and universal attacker, it is inexpensive and straightforward
to disconnect a user from these networks by blocking access to well known and
published resources. Once these avenues of attack have been exhausted it is
still possible to connect to the network by connecting to lesser known access
points. Since these networks are capable of communicating using many of the
same methods as most conventional Internet applications, a blanket ban on this
kind of communication would be too broad and extensive.

Tor employs strong encryption, the goal of which is to ensure that the data
transmitted is indistinguishable from random noise. To an observer, data
transmitted using encryption looks the same. Thus the traditional means of
identifying these protocols are ineffective.

This raises the question: is it possible to distinguish the traffic created by
these networks, from encrypted traffic produced by conventional software? This
question can be broken down further:

\begin{enumerate*}
  \item Can the traffic be classified using automated matching techniques?
  \item Do these networks have characteristics that make them readily
    distinguishable using heuristics based matching?
  \item Do they have characteristics that make them distinguishable using a
    machine learning technique?
  \item How long does a user have to be
    connected to the network before a confident match can be made?
\end{enumerate*}

@@

Research into Tor weaknesses has included the use of hostile exit nodes to
capture private information \parencite{website:tor-password-leak}, attacks that
take advantages of web based technologies such as Javascript and Flash to
reveal user's identities \parencite{Abbott:2007p298} and a number of identity
revealing traffic analysis techniques
\parencite{Murdoch:2005p325,Abbott:2007p298,Evans:2009p315}. Notable is the
discovery of a method for identifying bridge operators as discussed in
\textcite{McLachlan:2009p197}. This attack takes advantages in weaknesses of
the Tor protocol to enumerate bridges and some approaches to mitigation are
suggested. It is believed that this is the first proposal of a traffic analysis
technique for classification of Tor network streams.

@@

\chapter{Theoretical Framework}

\section{Research Methodology}

The experiment conducted in this paper utilize a non-equivalent group design.
The experimental groups are generated using simulations which allow for a great
deal of control over confounding variables, ensuring minimal difference between
the control and treatment groups. Both experiment groups perform the exact same
simulations. The control group performs all simulations using regular HTTPS,
while the treatment group proxies all communications over Tor. The treatment
group can be broken into HTTP over Tor and HTTPS over Tor based simulations.

\section{Assumptions}

It is assumed that the usage patterns exhibited by individual users will be
smaller than the communications characteristics that will lead to the
identification of anonymous and censorship resistant networks. Thus there is no
need to obtain a large sample of regular network traffic from varying user
profiles.

As of the current implementation, Tor network traffic is readily
distinguishable by looking at the handshake packets. It is likely that this
weakness will be addressed in a future version of the Tor protocol as it is
recognised as a design goal in \textcite{Dingledine:2008p1542}. For this
reason, this proposal focuses on traffic analysis techniques that are content
agnostic.

\section{Variables}

The data capturing stage will be affected by a number of variables that will
influence the accuracy of the chosen matching algorithms.

\subsection{Quality of the Anonymous Network}

An individual connection to the Tor network traverses a number of nodes to
create a circuit. The path a circuit uses is first defined by the exit node,
then chosen based on a number of constraints with a preference for high
bandwidth relays \parencite{website:tor-path-selection}. Since each connection
uses a different circuit and bandwidth demands are constantly changing, it is
difficult to predict the composition of a Tor circuit.

This means that any individual Tor circuit could be composed of systems with
differing capabilities, network throughput and configuration.

For testing purposes, the Tor network can be simulated on a small isolated
network. This will eliminate a lot of the variability that comes from
monitoring the real Tor network. Factors that affect the quality of the network
can be controlled by limiting the number of competing network applications,
controlling the number of running Tor instances and ensuring a consistent
configuration across all nodes.

\subsection{System Performance}

The ability of Tor to route packets in a timely fashion is highly dependent on
the performance of the system it is executing on. This is greatly compounded by
the usage of strong encryption which is computationally expensive. Packet
latency is one of the key parameters used in traffic classification techniques.
Performance related variables include:

\begin{itemize*}
  \item CPU speed and number of cores
  \item RAM speed and capacity
  \item Performance of integrated components, system bus etc.
  \item Hard disk size, access time and throughput
  \item Installed and running applications
  \item System configuration
\end{itemize*}

To remedy this, the systems used to simulate the Tor network should be made
uniform. This means an identical hardware and software configuration for all
systems in the simulation network.

In addition it is noted that system performance can degrade over time due to
both memory and hard disk fragmentation. This can be avoided by running the
simulations in a virtualisation environment that support snapshots, rolling
back to a known configuration point after each test.

\subsection{Network Performance}

Like system performance, latency is also dependent on the quality of the
network. Network switching equipment has limited resources for routing packets
and competing network traffic can introduce congestion which would radically
alter the observable communications patterns \parencite{Jacobson:1995p6768}.
Tor also supports congestion mitigation techniques which alter communication
patterns \parencite[8]{Dingledine:2004p314}.

Latency and network performance over the Internet are highly variable and
beyond the control of any one individual. Usage patterns vary according to
geographic region, societal conventions and the period in which they occur
\parencite{Thompson97wide-areainternet,Ken03longitudinalstudy}. The Internet,
and by necessity the public Tor network are globally distributed, which means
that circuits traverse a large variety of unpredictable network conditions. By
isolating the test network, these variables can be controlled ensuring low
latency high bandwidth communications or congestion as necessary.

\subsection{Application Protocol}

Tor provides communications facilities to a number of Internet enabled
applications by providing an interface using the SOCKS protocol
\parencite[17]{Dingledine:2004p314}. This means that the Tor network can proxy
regular web browsing, email or any other TCP or UDP based communication protocol
\parencite{website:socks}. Each of these application protocols have their own
distinguishing communications characteristics which will influence the matching
algorithm. It is not yet clear how transparent the Tor network will be regarding
these characteristics and thus, for the purposes of classification a single
protocol should be chosen.

\subsection{Caching}

Both web browsers and web servers are designed for high performance and often
cache requests so that they can be delivered faster in the future
\parencite{Caceres:1998p7419}. While this is a normal part of communications
traffic and should be considered, it is important that both regular and cached
requests appear uniformly across all experiments. Manually flushing the cache
or rolling back to a known virtual machine snapshot between each simulation can
ensure this outcome.

\input experiment.tex

\chapter{Results}

\section{Results}

The classification results are tabulated below. The best performing algorithm
j4.8 was able to classify HTTP over Tor with 97.8\% accuracy and a false
positive rate of 4.3\%. HTTPS was more easily identified with a 97\% accuracy
and low false positive rate of 0.7\%.

\begin{table}
  \begin{tabular}{lrrr}
    \toprule
    & True Positive Rate & False Positive Rate & ROC \\
    \midrule
    Random Forest\\
    \midrule
    HTTPS & 0.957 & 0.036\\
    HTTP over Tor & 0.937 & 0.037\\
    HTTPS over Tor & 0.977 & 0.003\\
    \midrule
    j4.8 With 10 fold cross validation\\
    \midrule
    HTTPS & 0.951 & 0.04\\
    HTTP over Tor & 0.978 & 0.043\\
    HTTPS over Tor & 0.97 & 0.007\\
    \midrule
    Random Tree\\
    \midrule
    HTTPS & 0.961 & 0.046\\
    HTTP over Tor & 0.906 & 0.04\\
    HTTPS over Tor & 0.955 & 0.01\\
    \midrule
    Adaboost\\
    \midrule
    HTTPS & 0.95 & 0.001\\
    HTTP over Tor & 0.999 & 0.324\\
    HTTPS over Tor & 0 & 0\\
    \bottomrule
    \label{table:results}
  \end{tabular}
  \caption{Results}
\end{table}

\chapter{Discussion of Results}

\section{Limitations}

The classifiers developed within this experiment have been trained against
a relatively small sample set of simulated data, which may only represent
a small fraction of the width and breadth of conditions available on a wider
area network. The characteristics that allow for classification of Tor network
may be masked by the natural variability introduced by a heterogeneous network.

\section{Addressing Research Questions}

\subsection{Can the traffic be classified using automated matching techniques?}

Yes, the classification algorithms trialled were all able to identify Tor traffic
with reasonable accuracy.

\subsection{Do these networks have characteristics that make them readily
distinguishable using heuristics based matching?}

\subsection{Do they have characteristics that make them distinguishable using a
machine learning technique?}

Yes.

\subsection{How long does a user have to be connected to the network before a
confident match can be made?}

\section{Possible Implications of Results}

The research results suggest that the strategies used by anonymous networks to
hide the identity of network participants, may make those anonymous networks
more easy to distinguish from networks and software that do not provide any
anonymity.

Research has shown that the low latency design of Tor means it is vulnerable to
traffic analysis attacks that correlate traffic patterns with users. This
vulnerability is indicative of the subtle amount of overhead incurred when
sending traffic through Tor. This should make it difficult to identify whether a
connection is using Tor VS not using Tor, though it is clearly still possible.
Reducing this overhead further may reduce the success of Tor traffic
identification.

Stronger anonymity providing networks utilize more sophisticated techniques to
hide individual identities such as batching, queuing and normalization. While
they may be more resistant to identity uncovering attacks, this may also make
them more easily identifiable by classification techniques.

\chapter{Conclusion}

Using Tor as a communications proxy incurs some overhead, sufficient enough that
when using a Tor proxy in controlled conditions, Tor traffic can be
distinguished from regularly encrypted traffic. This overhead may be sufficient
enough that Tor nodes in a real world network can be identified by using readily
available eavesdropping techniques.

However, the experiment was based on a small set of simulated data, with which
it would be impossible to cover all the possible real world conditions. The
variability present in the real Tor network may make this classification
technique impossible. Further research needs to be conducted with live packet
traces from real participants in the Tor network.

\chapter{Appendix}

\section{Manager Script}

\label{manager-script}
\lstinputlisting[language=ruby]{../experiment/simulation/manager/run_simulation.rb}

\section{Client Script}

\label{client-script}
\lstinputlisting[language=ruby]{../experiment/simulation/client/client.rb}

\section{Generate Tor Network Script}

\label{generate-tor-network-script}
\lstinputlisting[language=sh]{../experiment/simulation/tor/make_private_network.sh}

\section{Run Tor Network Script}

\label{run-tor-network-script}
\lstinputlisting[language=sh]{../experiment/simulation/tor/run_private_network.sh}

\printbibliography[title=REFERENCES]

\end{document}

% vim: fdm=syntax tw=80

>>>>

@@

Research into Tor weaknesses has included the use of hostile exit nodes to
capture private information \parencite{website:tor-password-leak}, attacks that
take advantages of web based technologies such as Javascript and Flash to reveal
user's identities \parencite{Abbott:2007p298} and a number of identity revealing
traffic analysis techniques
\parencite{Murdoch:2005p325,Abbott:2007p298,Evans:2009p315}.
\citeauthor{McLachlan:2009p197} reveals a method for identifying bridge
operators in \textcite{McLachlan:2009p197} using the Tor protocol directly.  It
is believed that this is the first proposal of a traffic analysis technique for
classification of Tor network streams.

@@
